---
title: "R Code for Lecture 8"
output: html_document
---

```{r, message = F, warning = F}
library(WindR)      
library(TSA)
library(MTS)
library(vars)
library(dplyr)
```

## 1. Daily Data of S&P 500, S&P 400 (MidCap) and S&P 600 (SmallCap) 

We download all data from "Wind Financial Terminal" (WTF) via its "Data Feed Services" (WDFS). To use WDFS in R/RSrtudio, one should first install the `WindR` package, then load it by `library(WindR)`[^1], and launch it using `w.start()`. 

[^1]: WDFS (WindR package) is not currently available for R/RStudio installed on Mac OS computers. 
```{r, message = F, warning = F, comment=FALSE, results=FALSE}
w.start()           # launch WindR DataFeed
  
index <- w.wsd("000001.SH,399106.SZ,HSCI.HI,SP500.SPI,SP400.SPI,SP600.SPI","close",
               "2011-01-01","2014-12-31","TradingCalendar=NYSE")$Data %>%
  rename(SH = `000001.SH`, SZ = `399106.SZ`, SP500 = `SP500.SPI`,
         SP400 = `SP400.SPI`, SP600 = `SP600.SPI`, date = `DATETIME`) %>%
  mutate(R.SH = (SH-lag(SH, n = 1L))/lag(SH, n = 1L),
         R.SZ = (SZ-lag(SZ, n = 1L))/lag(SZ, n = 1L),
         R.SP500 = (SP500-lag(SP500, n = 1L))/lag(SP500, n = 1L),
         R.SP400 = (SP400-lag(SP400, n = 1L))/lag(SP400, n = 1L),
         R.SP600 = (SP600-lag(SP600, n = 1L))/lag(SP600, n = 1L)) 
```

***

## 2. LSE of Vector Autoregressive Model using S&P 500/400/600 Data

```{r, message = F, warning = F}
data <- filter(select(index, R.SP500, R.SP400, R.SP600), !is.na(R.SP500))
```

### 2.1 Fitting a VAR($p$) model with $p=1$ selected by BIC 

Note: To use the `VAR()` function below, one need first install the `vars` package.
```{r, message = F, warning = F}
SPvar.bic <- VAR(data, lag.max = 3, ic = "SC")
summary(SPvar.bic)
```

```{r, message = F, warning = F}
## Refitting the model by leaving out insignificant coefficients
restrict(SPvar.bic, thresh = 1.9)
```

```{r, message = F, warning = F}
## ACF of residuals
ACF.SPvar.bic <- acf(residuals(SPvar.bic), na.action = na.pass, plot = F)
plot(ACF.SPvar.bic)
```

```{r, message = F, warning = F}
## Portmanteau tests
serial.test(SPvar.bic)
```

Note: Using AIC (set `ic = "AIC"`) and HQIC result in a VAR($p$) model with $p=3$ and $p=1$, respectively.
 
### 2.2 Fitting a VAR($p$) model with $p=2$ 
```{r, message = F, warning = F}
SPvar2 <- VAR(data, p = 2)
summary(SPvar2)
```

```{r, message = F, warning = F}
## Refitting the model by leaving out insignificant coefficients
restrict(SPvar2, thresh = 1.9)
```

```{r, message = F, warning = F}
## ACF of residuals
ACF.SPvar2 <- acf(residuals(SPvar2), na.action = na.pass, plot = F)
plot(ACF.SPvar2)
```

```{r, message = F, warning = F}
## Portmanteau tests
serial.test(SPvar2)
```

Note: The MLE of a VAR($p$) model can be implemented by the `VARMA()` function in the `MTS` package.

## 3. LSE of Vector Autoregressive Model using S&P500, SH and SZ Data

```{r, message = F, warning = F}
data <- filter(select(index, R.SP500, R.SH, R.SZ), !is.na(R.SP500*R.SH))
```

```{r, message = F, warning = F}
var.bic <- VAR(data, lag.max = 3, ic = "SC")
summary(var.bic)
## Refitting the model by leaving out insignificant coefficients
restrict(var.bic, thresh = 1.9)
ACF.var.bic <- acf(residuals(var.bic), na.action = na.pass, plot = F)
plot(ACF.var.bic)
serial.test(var.bic)
```

## Causality

```{r, message = F, warning = F, comment=F, results=F}
# Download data from Wind and clean it
w.start()           
index <- w.wsd("SP400.SPI,SP500.SPI","close",
               "2011-01-01","2014-12-31","TradingCalendar=NYSE")$Data %>%
  rename(SP400 = `SP400.SPI`, SP500 = `SP500.SPI`, date = `DATETIME`) %>%
  mutate(R.SP400 = (SP400-lag(SP400, n = 1L))/lag(SP400, n = 1L),
         R.SP500 = (SP500-lag(SP500, n = 1L))/lag(SP500, n = 1L)) 
```

```{r, message = F, warning = F}
# LSE of VAR(2) model of S&P 500/400 index
data <- filter(select(index, R.SP400, R.SP500), !is.na(R.SP500*R.SP400))
var.lse <- VAR(data, ic = "SC")
summary(var.lse)
```

```{r, message = F, warning = F}
# Test if S&P 400 index Granger causes S&P 500 index
causality(var.lse, cause = "R.SP400")
# Test if S&P 500 index Granger causes S&P 400 index
causality(var.lse, cause = "R.SP500")
```

```{r, message = F, warning = F}
var.lse <- VAR(data, p = 2)
summary(var.lse)
```

```{r, message = F, warning = F}
Psi(var.lse)
irf.var.lse <- irf(var.lse)
plot(irf.var.lse)
```

## ECM

```{r, message = F, warning = F}
# Monte Carlo simulation of ECM model
n <- 200; e <- rnorm(n)
X = cumsum(e)
Y = X + rnorm(n, 0, 1/2)

t <- lm(Y ~ X)
hatZ <- residuals(t)[-c(1, n)]
summary(t)

dX <- diff(X); dY <- diff(Y)
data <- data.frame(embed(cbind(dY, dX), 2), hatZ)
colnames(data) <- c("dY0", "dX0", "dY1", "dX1", "hatZ")
attach(data)

ecm <- lm(dY0 ~ hatZ + dX1)
summary(ecm)

# Use ECMvar

```


